Feature importance analysis

### 2020-02-28 analysis/feature-importance.Rmd
save feature-importance state

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.retina = 3,
  fig.align = "center",
  fig.width = 6.93,
  fig.height = 6.13,
  out.width = "100%",
  echo = FALSE
)

library("dplyr")
library("tidyr")
library("ggpubr")
library("purrr")
library("glue")
library("fs")
R.utils::sourceDirectory("R")

# load drake objects
drake::loadd(
)

library(dplyr)
library(hsdar)
```

```{r }
# create datasets of most imp features for each feature set
data_vi = data.frame(class = 
                    c("VI", "VI", "VI"),
                  rank = 
                    c(1, 3, 3),
                  wavelength = 
                    c(404, 404, 788),
                  feature = c("FeatureA", "FeatureB", "FeatureC")
)

data_nri = data.frame(class = 
                    c("NRI", "NRI", "NRI"),
                  rank = 
                    c(1, 3, 3),
                  wavelength = 
                    c(4050, 470, 600),
                  feature = c("FeatureA", "FeatureB", "FeatureC")
)

# create spectral signature of vegetation 
# PROSAIL is a algorithm simulating spectral signatures of vegetation, see `?hsdar::PROSAIL`
# reflectance is scaled to 0-10 to be able to plot it in the same plot as the filter rankings -> the y-axes need to be the same
# PROSAIL returns a spectral signature from 400 nm to 2500 nm -> we take the values only and subset to 400 nm to 1000 nm
# because we order from 1 - 10 with 1 being the best, we have to reverse the scaling of the reflectance values
spectra_sim = hsdar::PROSAIL()
spectra_df = data.frame(reflectance = as.vector(spectra_sim@spectra@spectra_ma),
                wavelength = seq(400, 2500, 1)) %>% 
  dplyr::filter(wavelength < 1000) %>% 
  # scale the reflectance to fit into the range of the y-axis for the filter ranking (the 10 - is to reverse the scale)
  dplyr::mutate(reflectance = 10 - scale(reflectance, center = FALSE, scale = max(reflectance, na.rm = TRUE)/10))

# to be able to plot the sec y axis on the plot, we need to supply the scaled R object -> we need to save the object separately
reflectance = as.vector(spectra_sim@spectra@spectra_ma)
reflectance_scaled = scale(reflectance, center = FALSE, scale = max(reflectance, na.rm = TRUE)/10)

# here we bind the simulated data with the filter rankings
# Important: Append the feature set class to the simulated data so when merged all together the facetting works (only works if we have class levels for all entries)
# This is also the reason why we cannot create an initial data.frame containing the results of all feature sets
data_vi_merged = bind_rows(spectra_df, data_vi) %>% 
  mutate(class = "VI")

data_nri_merged = bind_rows(spectra_df, data_nri) %>% 
  mutate(class = "NRI")

data_all = bind_rows(data_vi_merged, data_nri_merged)

ggscatter(data_all, x = "wavelength", y = "rank", color = "class", 
          #add = "segments",
          palette = "nejm",
          size = 3,
          facet.by = "class"
) +
  geom_segment(aes(x = wavelength, y = 10, xend = wavelength, yend = rank), color = "grey") + 
  #scale_y_continuous(limits = c(0L, 10L), breaks = scales::pretty_breaks()) +
  scale_x_continuous(limits = c(400, 1000), breaks = scales::pretty_breaks()) +
  scale_y_reverse(limits = c(10L, 0L), breaks = scales::pretty_breaks(), sec.axis = sec_axis(~ scale(-., center = FALSE, scale = max(., na.rm = TRUE)/-1 ),
                                                                                             labels = c(10, 0.75, 0.55, 0.25, 0),
                                                                                             name = "Reflectance [%]")) +
  labs(y = "Importance", x = "Wavelength [nm]", color = "Feature set") +
  geom_line(aes(x = wavelength, y = reflectance), linetype = "dotted") +
  ggrepel::geom_label_repel(label = data_all$feature, nudge_x = 0.5, nudge_y = 0.5)
```

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.retina = 3,
  fig.align = "center",
  fig.width = 8.5,
  fig.asp = 0.66,
  out.width = "100%",
  echo = FALSE
)

library("drake")
library("hsdar")
library("dplyr")
library("ggplot2")
library("ggpubr")
library("ggpmisc")
library("patchwork")

# load drake objects
loadd(
  fi_permut_hr_buffer2,
  fi_permut_vi_buffer2
)
```

Sort the permutation feature importance results

```{r feature-importance-1 }
fi_ranked_hr <- fi_permut_hr_buffer2$res %>%
  tibble::rownames_to_column("measure") %>%
  tidyr::pivot_longer(
    cols = starts_with("B"),
    values_to = "importance", names_to = "feature"
  ) %>%
  mutate(wavelength = seq(420, 995, 4.75)) %>%
  mutate(numeric_id = seq(5, 126, 1)) %>%
  arrange(desc(importance)) %>%
  mutate(rank = row_number()) %>%
  select(-measure)
fi_ranked_hr

fi_ranked_vi <- fi_permut_vi_buffer2$res %>%
  tibble::rownames_to_column("measure") %>%
  tidyr::pivot_longer(
    cols = starts_with("B"),
    values_to = "importance", names_to = "feature"
  ) %>%
  # mutate(wavelength = seq(420, 995, 4.75)) %>%
  arrange(desc(importance)) %>%
  mutate(rank = row_number()) %>%
  select(-measure)
fi_ranked_vi
```

# Create a virtual spectral signature of vegetation using PROSAIL.

PROSAIL is a algorithm simulating spectral signatures of vegetation, see `?hsdar::PROSAIL`.
Reflectance is scaled to 0-10 to be able to plot it in the same plot as the feature importance rankings -> the axis limits for the y and z axis needs to match.

PROSAIL returns a spectral signature from 400 nm to 2500 nm -> we take the values only and subset to 400 nm - 1000 nm.
Because we order from 1 - 10 with 1 being the best rank, we have to reverse the scaling of the reflectance values.

```{r feature-importance-2 }
spectra_sim <- hsdar::PROSAIL()
spectra_df <- data.frame(
  reflectance = as.vector(spectra_sim@spectra@spectra_ma),
  wavelength = seq(400, 2500, 1)
) %>%
  dplyr::filter(wavelength < 1000) %>%
  # scale the reflectance to fit into the range of the y-axis for the filter
  # ranking (the 10 - is to reverse the scale)
  dplyr::mutate(reflectance = 10 - scale(reflectance,
    center = FALSE,
    scale = max(reflectance, na.rm = TRUE) / 10
  ))

# to be able to plot the sec y axis on the plot, we need to supply the scaled R
# object -> we need to save the object separately
reflectance <- as.vector(spectra_sim@spectra@spectra_ma)
reflectance_scaled <- scale(reflectance,
  center = FALSE,
  scale = max(reflectance, na.rm = TRUE) / 10
)
```

Next we bind the simulated data with the feature importance rankings.
To join both data.frames we need to round the reflectance centers of the bands to integers to match with the reflectance values created by PROSAIL.

```{r feature-importance-3 }
fi_ranked_hr$wavelength <- round(fi_ranked_hr$wavelength)

data_hr_merged <- left_join(spectra_df, fi_ranked_hr, by = c("wavelength")) %>%
  mutate(class = "HR")

# data_vi_merged <- left_join(spectra_df, fi_ranked_vi, by = c("wavelength")) %>%
#   mutate(class = "vi") %>%
#   mutate(reflectance = as.numeric(.data$reflectance))
```

```{r feature-importance-4, echo=FALSE}
plot_reflectance_imp_rank <- function(data) {
  ggplot(data, aes(x = .data[["wavelength"]], y = .data[["rank"]])) +
    geom_line(aes(x = wavelength, y = reflectance, color = "Spectral Signature of Vegetation"),
      linetype = "solid",
      size = 0.3
    ) +
    geom_point(size = 2, aes(color = "Band"), show.legend = T) +
    scale_x_continuous(limits = c(400, 1000), breaks = scales::pretty_breaks()) +
    scale_y_reverse(
      expand = c(0, 0),
      limits = c(10L, 0),
      # breaks = scales::pretty_breaks(),
      breaks = seq(1, 10, 1),
      sec.axis = sec_axis(~ scale(-.,
        center = FALSE,
        scale = max(., na.rm = TRUE) / -1
      ),
      labels = c(1.0, 0.75, 0.55, 0.25, 0),
      name = "Reflectance [%]"
      )
    ) +
    guides(color = guide_legend(
      title = NULL,
      override.aes = list(
        linetype = c("blank", "solid", "solid"),
        shape = c(19, NA, NA)
      )
    )) +
    labs(
      title = "Permutation-based Variable Importance Results",
      subtitle = paste0(
        "Top ten features of 'VI' and 'HR' datasets along a simulated spectral signature of vegetation.\n",
        "Labels show the hyperspectral band number of dataset 'HR'."
      ),
      caption = "Learner: SVM; 100 Monte-Carlo Iterations",
      y = "Importance Rank", x = "Wavelength [nm]"
    ) +
    ggrepel::geom_label_repel(
      label = data$feature,
      nudge_x = 0.5, nudge_y = 0.5,
      label.size = 0.15
    ) +
    ggpubr::theme_pubclean()
}

plot_reflectance_imp_absolute <- function(data, x_identifier, class) {
  pl <- ggplot(data, aes(x = .data[[x_identifier]], y = .data[["importance"]])) +
    labs(y = "Importance", x = "Band number") +
    geom_segment(aes(
      x = .data[[x_identifier]], y = 0,
      xend = .data[[x_identifier]], yend = .data[["importance"]]
    ),
    color = "grey", show.legend = FALSE
    ) +
    geom_point(size = 1, color = "black", show.legend = T) +
    labs(
      title = glue::glue("Permutation-based Variable Importance for Dataset '{class}'"),
      subtitle = "Absolute importance values by band",
      caption = "Learner: SVM; 100 Monte-Carlo iterations"
    ) +
    ggpubr::theme_pubclean()

  if (is.character(data[[x_identifier]])) {
    pl
  } else {
    pl + scale_x_continuous(breaks = seq(5, 125, 5))
    pl
  }
}
```

# Plots by dataset

## HR

### P1 Absolute permutation based Var Imp

```{r fi-permut-hr-abs, dev = c("png", "pdf")}
p1 <- fi_ranked_hr %>%
  plot_reflectance_imp_absolute("numeric_id", class = "HR")
p1
```

### P2 Top ten features along the spectral curve

```{r feature-importance-6}
table_inds <- tibble::tribble(
  ~Rank, ~Name, ~Formula,

  "1", "Vogelmann2", "(R_{734}-R_{747})/(R_{715}+R_{726})",
  "2", "Vogelmann4", "(R_{734}-R_{747})/(R_{715}+R_{720})",
  "3", "Vogelmann", "R_{740}/R_{720}",
  "4", "NPCI", "(R_{680}-R_{430})/(R_{680}+R_{430})",
  "5", "Vogelmann3", "D_{715}/D_{705}",
  # "D2"        , "$D_{705}/D_{722}$",
  # "Datt3"     , "$D_{754}/D_{704}$",
  # "PWI"       , "$R_{900}/R_{970}$",
  # "SR7"       , "$R_{440}/R_{690}$",
  # "SRPI"      , "$R_{430}/R_{680}$"
)
```

```{r fi-permut-vi-hr-ranked, dev = c("png", "pdf"), warning=FALSE}
table_inset <- tibble(x = 0.4, y = 1.15, tb = list(table_inds))

p2 <- data_hr_merged %>%
  plot_reflectance_imp_rank() +
  # Vogelmann 2 (P1)
  geom_segment(aes(
    x = 715, xend = 747,
    y = 1, yend = 1,
    color = "Range of VegIndex"
  ), size = 0.5) +

  # Vogelmann 4 (P2)
  geom_segment(aes(
    x = 720, xend = 747,
    y = 2, yend = 2
  ), color = "green", size = 0.5) +

  # Vogelmann  (P3)
  geom_segment(aes(
    x = 720, xend = 740,
    y = 3, yend = 3
  ), color = "green", size = 0.5) +

  # NPCI (P4)
  geom_segment(aes(
    x = 680, xend = 690,
    y = 4, yend = 4
  ), color = "green", size = 0.5) +
  geom_segment(aes(
    x = 430, xend = 440,
    y = 4, yend = 4
  ), color = "green", size = 0.5) +

  # Vogelmann 3 (P5)
  geom_segment(aes(
    x = 705, xend = 715,
    y = 5, yend = 5,
  ), color = "green", size = 0.5) +

  # D2 (P6)
  geom_segment(aes(
    x = 705, xend = 722,
    y = 6, yend = 6
  ), color = "green", size = 0.5) +

  # Datt3 (P7)
  geom_segment(aes(
    x = 704, xend = 709,
    y = 7, yend = 7
  ), color = "green", size = 0.5) +
  geom_segment(aes(
    x = 754, xend = 759,
    y = 7, yend = 7
  ), color = "green", size = 0.5) +

  # PWI (P8)
  geom_segment(aes(
    x = 900, xend = 910,
    y = 8, yend = 8
  ), color = "green", size = 0.5) +
  geom_segment(aes(
    x = 970, xend = 980,
    y = 8, yend = 8
  ), color = "green", size = 0.5) +

  # SR7 (P9)
  geom_segment(aes(
    x = 440, xend = 450,
    y = 9, yend = 9
  ), color = "green", size = 0.5) +
  geom_segment(aes(
    x = 690, xend = 700,
    y = 9, yend = 9
  ), color = "green", size = 0.5) +

  # SRPI (P10)
  geom_segment(aes(
    x = 430, xend = 440,
    y = 10, yend = 10
  ), color = "green", size = 0.5) +
  geom_segment(aes(
    x = 680, xend = 690,
    y = 10, yend = 10
  ), color = "green", size = 0.5) +
  # annotate(geom = "table", x = 395, y = 0,
  #          label = list(table_inds), size = 2.4,
  #        vjust = 1, hjust = 0)
  ggpmisc::geom_table_npc(
    data = table_inset, aes(npcx = x, npcy = y, label = tb),
    hjust = 1, vjust = 1, size = 2.35
  ) +
  scale_color_manual(
    values = c(
      "Band" = "black", "Spectral Signature of Vegetation" = "black",
      "Range of VegIndex" = "green"
    )
  )

p2
```

Combined

```{r feature-importance-8, fig.width=9.5, fig.asp=1.1}
p1 / p2
```

## VI

### P3 Absolute permutation based Var Imp

```{r fi-permut-vi-abs, dev = c("png", "pdf")}
p3 <- fi_ranked_vi %>%
  plot_reflectance_imp_absolute("feature", class = "VI") +
  ggpubr::rotate_x_text()
p3
```

Vogelmann2 $(R_{734}-R_{747})/(R_{715}+R_{726})$ Vogelmann et al. (1993)
Vogelmann4 $(R_{734}-R_{747})/(R_{715}+R_{720})$ Vogelmann et al. (1993)
Vogelmann3 $D_{715}/D_{705}$ Vogelmann et al. (1993)
Vogelmann  $R_{740}/R_{720}$ Vogelmann et al. (1993)
NPCI       $(R_{680}-R_{430})/(R_{680}+R_{430})$
D2         $D_{705}/D_{722}$
Datt3      $D_{754}/D_{704}$
PWI        $R_{900}/R_{970}$
SR7        $R_{440}/R_{690}$
SRPI       $R_{430}/R_{680}$

Dxxx: First derivation of reflectance values at wavelength 'xxx'.
Rxxx: Reflectance at wavelength 'xxx'.

Reference: `?hsdar::vegindex()`
